Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-09-27 saner/paper.tex
Updates from Overleaf

\documentclass[]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{amsmath}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

@@

We analyze the overhead of applying a refactoring in terms of four factors: the size of the clone, the relation between the code fragments in a clone, whether clone fragments create, modify or returns data and the amount of external data that cloned fragments use. We find that the combined volume of similar fragments in a clone is the biggest influencing factor: the majority of duplicates with a total size of 63 or more tokens improve maintainability when refactored. The amount of external data that needs to be passed to the merged location of the duplicate is the other important factor: the majority of clones with more than requiring more than one external parameter decreases maintainability.
\end{abstract}

\begin{IEEEkeywords}
code clones, refactoring, static code analysis, object-oriented programming
\end{IEEEkeywords}

@@

@@

\section{The CloneRefactor tool}
To detect such clones that can be refactored, we surveyed a set of modern clone detection tools and techniques~\cite{svajlenko2014evaluating, sheneamer2016survey, gautam2016various, roy2009comparison}. We created a set of control projects to determine the suitability of these tools to detect refactorable clones, either through configuration or postprocessing of their output.

None of the surveyed tools~\cite{ragkhitwetsagul2019siamese, roy2008nicad, semura2017ccfindersw, sajnani2016sourcerercc, saini2018oreo, jiang2007deckard, kamalpriya2017enhancing, mazinanian2016jdeodorant} were suitable with our findings about refactorable clones, because of which we propose CloneRefactor. This tool goes through a 3-step process to automatically refactor clones as shown in Fig.~\ref{fig:clonerefactorflow}. In the following sections, we explain these steps.
\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{img/flow}
  \caption{Simple flow diagram of CloneRefactor.}
  \label{fig:clonerefactorflow}
\end{figure}

@@

@@

CloneRefactor requires all libraries of the projects it analyses, to find the full paths of all referenced symbols in the source code (see Sec.~\ref{sec:t1r}). Because of that requirement, we decided to filter the corpus to only projects using the Maven build automation system. We created a set of scripts to prepare such a corpus with all dependencies included.

@@

\subsection{Tool Validation}
We have validated the correctness of CloneRefactor through unit tests and empirical validation. First, we created a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, and refactorability category to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context, and their proposed refactoring.

@@

\subsection{Thresholds}
In our corpus, CloneRefactor has refactored 12.710 clone classes and measured the change in indicated metrics (see Sec.~\ref{sec:metrics}). Using the presented formulas (see Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

\subsubsection{Clone Token Volume}
TODO shows the obtained results when plotting the clone volume vs the average delta maintainability score. We define the \textit{token volume} as the combined number of tokens in all clone instances in a refactored clone class. For higher token volume numbers we have fewer refactorings that refactor such clones, because of which we represent the x-axis as a logarithmic scale. The trendline intersects the ``zero'' line (maintainability does not increase nor decrease) at a token volume of 63.

@@

\subsubsection{Parameters}
TODO shows how an increase in parameters lowers the maintainability of the refactored code. On the primary x-axis, the maintainability is displayed. The secondary x-axis shows the number of refactorings. The y-axis shows the number of parameters.

@@

\bibliographystyle{IEEEtran}
\bibliography{references}

>>>>

\documentclass[sigconf,review, table]{acmart}

\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{float}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{parcolumns}

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{pred}{rgb}{0.9,0,0}
\definecolor{pgrey}{rgb}{0.46,0.45,0.48}

\usepackage{listings}
\lstset{language=Java,
  showspaces=false,
  showtabs=false,
  breaklines=true,
  showstringspaces=false,
  breakatwhitespace=true,
  commentstyle=\color{pgreen},
  keywordstyle=\color{pblue},
  numbers = left,
  stringstyle=\color{pred},
  basicstyle=\ttfamily,
  moredelim=[il][\textcolor{pgrey}]{\$\$},
  moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%}
}

\definecolor{lightyellow}{HTML}{fffdcc}
\definecolor{doublelightyellow}{HTML}{fffa9c}
\definecolor{lightgreen}{HTML}{dcffd4}
\definecolor{lightblue}{HTML}{e0ffff}
\definecolor{lightred}{HTML}{ffe0e0}
\definecolor{lightpink}{HTML}{ffe0ff}
\definecolor{lightorange}{HTML}{ffede0}
\definecolor{lightpurple}{HTML}{f0e0ff}
\newcommand{\highlightYellow}{\makebox[0pt][l]{\color{lightyellow}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightGreen}{\makebox[0pt][l]{\color{lightgreen}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightBlue}{\makebox[0pt][l]{\color{lightblue}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightRed}{\makebox[0pt][l]{\color{lightred}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightPink}{\makebox[0pt][l]{\color{lightpink}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightOrange}{\makebox[0pt][l]{\color{lightorange}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightPurple}{\makebox[0pt][l]{\color{lightpurple}\rule[-4pt]{1\linewidth}{14pt}}}
\newcommand{\highlightDarkyellow}{\makebox[0pt][l]{\color{doublelightyellow}\rule[-4pt]{1\linewidth}{14pt}}}

\lstnewenvironment{javacode}[1][]{\lstset{language=Java,escapechar=|,tabsize=2, breaklines=true, xleftmargin=.25in, keywordstyle=\color{pblue},basicstyle=\small,#1}}{}

\setcopyright{acmcopyright}
\copyrightyear{2019}
\acmYear{2019}
\acmDOI{10.1145/1122445.1122456}
\acmMonth{8}

\acmConference[ICSE 2020]{Technical Papers}{May 23--29}{Seoul, South Korea}

\begin{document}

@@

%\author{Simon Baars}
%\affiliation{\institution{University of Amsterdam}}
%\email{simon.mailadres@gmail.com}

%\author{Ana Oprescu}
%\affiliation{\institution{University of Amsterdam}}
%\email{a.m.oprescu@uva.nl}

@@

We analyze the overhead of applying a refactoring in terms of four factors: the size of the clone, the relation between the code fragments in a clone, whether clone fragments create, modify or return data and the amount of external data that cloned fragments use. We find that the combined volume of similar fragments in a clone is the biggest influencing factor: the majority of duplicates with a total size of 63 or more tokens improve maintainability when refactored. The amount of external data that needs to be passed to the merged location of the duplicate is the other important factor: the majority of clones with more than requiring more than one external parameter decreases maintainability.
\end{abstract}

\keywords{code clones, refactoring, static code analysis, object-oriented programming}

@@

\begin{figure*}
\begin{parcolumns}{2}
\colchunk[1]{
\begin{javacode}
package com.sb.game;

import java.util.List;

public class GameScene
{
|\highlightYellow|	public void addToList(List l) {
|\highlightYellow|		l.add(getClass().getName());
|\highlightYellow|	}

	public void addTen(int x) {
|\highlightYellow|		x = x + 10; // Addition
|\highlightYellow|		Notifier.notifyChanged(x);
|\highlightYellow|		return x;
	}
}
\end{javacode}}
\colchunk[2]{
\begin{javacode}
package com.sb.fruitgame;

import java.awt.List; // Different type

public class LoseScene
{
|\highlightYellow|	public void addToList(List l) {
|\highlightYellow|		l.add(getClass().getName());
|\highlightYellow|	}

	public void concatTen(String x) {
|\highlightYellow|		x = x + 10; // String concat
|\highlightYellow|		Notifier.notifyChanged(x);
|\highlightYellow|		return x;
	}
}
\end{javacode}}
\end{parcolumns}
\caption{Example of textually equal code fragments that differ in functionality.}
\label{fig:type1}
\end{figure*}

@@

\section{The CloneRefactor tool}
To detect such clones that can be refactored, we surveyed a set of modern clone detection tools and techniques~\cite{svajlenko2014evaluating, sheneamer2016survey, gautam2016various, roy2009comparison}. We created a set of control projects\footnote{Control projects for assessing clone detection tools: \url{https://github.com/SimonBaars/CloneRefactor/tree/master/src/test/resources}} to determine the suitability of these tools to detect refactorable clones, either through configuration or postprocessing of their output.

None of the surveyed tools~\cite{ragkhitwetsagul2019siamese, roy2008nicad, semura2017ccfindersw, sajnani2016sourcerercc, saini2018oreo, jiang2007deckard, kamalpriya2017enhancing, mazinanian2016jdeodorant} were suitable with our findings about refactorable clones, because of which we propose CloneRefactor\footnote{The source code of CloneRefactor is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}}. This tool goes through a 3-step process to automatically refactor clones as shown in Fig.~\ref{fig:clonerefactorflow}. In the following sections, we explain these steps.
\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{img/flow}
  \caption{Simple flow diagram of CloneRefactor.}
  \label{fig:clonerefactorflow}
\end{figure}

@@

\subsubsection{Characteristics of the extracted method}\label{sec:characteristics}
We define the following characteristics of the extracted method and/or the call:
\begin{itemize}
\item \textbf{Tokens}: The number of tokens in the body of the method.
\item \textbf{Relation}: The relation category (Sec.~\ref{sec:relation}) by which this methods' location was determined.
\item \textbf{Returns}: Whether the method calls return, declare, assign or don't use any data from the extracted method.
\item \textbf{Parameters}: The number of parameters the extracted method has.
\end{itemize}
We hypothesize that these characteristics are the main factors influencing the impact on the maintainability of the system as a result of refactoring the clone.

@@

CloneRefactor requires all libraries of the projects it analyses, to find the full paths of all referenced symbols in the source code (see Sec.~\ref{sec:t1r}). Because of that requirement, we decided to filter the corpus to only projects using the Maven build automation system. We created a set of scripts\footnote{All scripts to prepare the corpus are available on GitHub: \url{https://github.com/SimonBaars/GitHub-Java-Corpus-Scripts}} to prepare such a corpus with all dependencies included.

@@

\subsection{Tool Validation}
We have verified the correctness of CloneRefactor through unit tests and empirical verification.\todo{With unit test you are testing the properties which is verification. With projects you are validating as you are checking whether its finding what is it supposed to be finding.} First, we created a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, and refactorability category to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context, and their proposed refactoring.

@@

\subsection{Thresholds}
In our corpus, CloneRefactor has refactored 12.710 clone classes and measured the change in indicated metrics (see Sec.~\ref{sec:metrics}). Using the presented formulas (see Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

\subsubsection{Clone Token Volume}
Figure \ref{fig:maintainabilityscore} shows the obtained results when plotting the clone volume vs the average delta maintainability score. We define the \textit{token volume} as the combined number of tokens in all clone instances in a refactored clone class. For higher token volume numbers we have fewer refactorings that refactor such clones, because of which we represent the x-axis as a logarithmic scale. The trendline intersects the ``zero'' line (maintainability does not increase nor decrease) at a token volume of 63.

\begin{figure*}
  \includegraphics[width=1\textwidth]{img/tokenvolume}
  \caption{A graph that shows how the size in tokens of the refactored clone affects maintainability.}
  \label{fig:maintainabilityscore}
\end{figure*}

@@

\subsubsection{Return Value}
Table~\ref{tab:return} shows how the return value of the extracted method influences the maintainability of the resulting system.

\begin{table*}
\centering
\begin{tabular}{@{}lrrrrrr@{}}
\toprule
\textit{\textbf{Return Category}} & \textit{\textbf{Complexity}} & \textit{\textbf{Parameters}} & \textit{\textbf{Size}} & \textit{\textbf{Duplication}} & \textit{\textbf{\#}} & \textit{\textbf{Score}} \\ \midrule
Return & 0.85 & 1.02 & -3.84 & -55.00 & 1,571 & 0.19 \\
Declare & 0.94 & 0.74 & 11.11 & -49.19 & 5,177 & 0.15 \\
\rowcolor[HTML]{FFCCC9}
Assign & 0.79 & 1.07 & 0.43 & -56.29 & 14 & 0.12 \\
Void & 0.76 & 1.49 & -5.85 & -56.35 & 5,921 & -0.18 \\ \midrule
\textbf{Grand Total} & \textbf{0.84} & \textbf{1.12} & \textbf{1.33} & \textbf{-53.26} & \textbf{12,683} & \textbf{0.00} \\ \bottomrule
\end{tabular}
\caption{Average metric values for refactorings of clone classes with the specified return category}
\label{tab:return}
\end{table*}

The displayed return categories are ordered by their scores. Overall, the scores do not deviate much (-0.18 to 0.19), indicating that the return category has a minor impact on maintainability. When the result of the call to the extracted method is directly returned, the maintainability score is the highest. When no value is returned, the maintainability score is the lowest. The main reason that no return value ends up lowest, is that it is linked to a higher number of parameters required for the extracted method.

\subsubsection{Parameters}
Fig.~\ref{fig:arguments} shows how an increase in parameters lowers the maintainability of the refactored code. On the primary x-axis, the maintainability is displayed. The secondary x-axis shows the number of refactorings. The y-axis shows the number of parameters.

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{img/arguments}
  \caption{Influence of number of method parameters on system maintainability.}
  \label{fig:arguments}
\end{figure}

@@

\section*{Acknowledgments}
We would like to thank the Software Improvement Group for their continuous support in this project. In particular, we would like to thank Xander Schrijen for his invaluable input in this research. Furthermore, we would like to thank Sander Meester for his proofreading efforts and feedback.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}
